---
title: "ARIMA"
output: html_document
date: "2023-10-06"
---

```{r}

library(tseries)
library(ggplot2)

df <- read.csv("./Data/BoxOffice_DJ.csv")
head(df)
```

```{r}

df <- df[, c('weekend' , 'top10_gross' , 'closing_price')]
head(df)
```

```{r}
# Scatter plot whatever

# Create a scatter plot with ggplot and blue points
ggplot(data = df, aes(x = closing_price, y = top10_gross)) +
  geom_point(color = "blue") +  # This changes the point color to blue
  labs(x = "Closing Price", y = "Top 10 Gross Revenue", title = "Scatter Plot of Top 10 Gross Revenue vs Closing Price") +
  theme_minimal()

```

```{r}
library(forecast)
order <- c(4, 1, 1)

# Fit the ARIMAX model
# Note: 'xreg' is the argument where you can specify the exogenous regressor
arimax_model <- Arima(df$top10_gross, order=order, xreg=df$closing_price)

# Output the summary of the model to check coefficients and model fit
summary(arimax_model)


```

```{r}
# Using auto arima 

fixed <- rep(NA, 6)  # for ARIMA(4,1,1) there are 4 AR, 0 differencing, and 1 MA parameters
fixed[c(1, 6)] <- 0  # This sets the first AR and the first MA parameters to 0, you adjust as needed

auto_arimax_model <- auto.arima(df$top10_gross, xreg=df$closing_price, d=1, start.p=4, start.q=1, max.p=4, max.q=1, max.order=NA, fixed=fixed, seasonal=FALSE)
summary(auto_arimax_model)
```

```{r}


library(forecast)
library(ggplot2)

# Assuming you have already divided your data into training and test sets
# Let's assume the last few observations are held out for testing
n_test <- 10  # Number of observations to hold out for testing
training_set <- head(df, -n_test)
test_set <- tail(df, n_test)

# Fit the manually specified ARIMAX model
arimax_model <- Arima(training_set$top10_gross, order=order, xreg=training_set$closing_price)

# Fit the auto.arima model
auto_arimax_model <- auto.arima(training_set$top10_gross, xreg=training_set$closing_price, d=1, start.p=4, start.q=1, max.p=4, max.q=1, max.order=NA, fixed=fixed, seasonal=FALSE)

# Generate predictions for the test set
arimax_forecast <- forecast(arimax_model, xreg=test_set$closing_price, h=n_test)
auto_arimax_forecast <- forecast(auto_arimax_model, xreg=test_set$closing_price, h=n_test)

# Calculate RMSE for each model
arimax_rmse <- sqrt(mean((test_set$top10_gross - arimax_forecast$mean)^2))
auto_arimax_rmse <- sqrt(mean((test_set$top10_gross - auto_arimax_forecast$mean)^2))

# Create a data frame for plotting
rmse_data <- data.frame(Model=c("ARIMAX", "AUTO ARIMAX"), RMSE=c(arimax_rmse, auto_arimax_rmse))

# Plot the RMSE for both models
ggplot(rmse_data, aes(x=Model, y=RMSE, fill=Model)) +
  geom_bar(stat="identity", position=position_dodge()) +
  labs(title="Comparison of RMSE between ARIMAX models", x="Model", y="RMSE") +
  theme_minimal()


```

```{r}
library(forecast)

# Define the total number of observations
n <- nrow(df)

# Number of folds for cross-validation
n_folds <- 5
fold_size <- floor(n / n_folds)

# Store the RMSE for each fold
rmse_arimax <- numeric(n_folds)
rmse_auto_arimax <- numeric(n_folds)

for (i in 1:n_folds) {
  # Define the training and testing indices
  test_indices <- ((i - 1) * fold_size + 1):(i * fold_size)
  if (i == n_folds) {
    # Ensure the last fold includes any leftover observations
    test_indices <- ((i - 1) * fold_size + 1):n
  }
  
  training_indices <- setdiff(1:n, test_indices)
  
  # Split the data into training and testing sets
  training_set <- df[training_indices, ]
  test_set <- df[test_indices, ]

  # Manually specified ARIMAX model
  arimax_model <- Arima(training_set$top10_gross, order=c(4, 1, 1), xreg=training_set$closing_price)
  arimax_forecast <- forecast(arimax_model, xreg=test_set$closing_price, h=length(test_indices))
  rmse_arimax[i] <- sqrt(mean((test_set$top10_gross - arimax_forecast$mean)^2))

  # AUTO ARIMA model
  auto_arimax_model <- auto.arima(training_set$top10_gross, xreg=training_set$closing_price, d=1, start.p=4, start.q=1, max.p=4, max.q=1, max.order=NA, seasonal=FALSE)
  auto_arimax_forecast <- forecast(auto_arimax_model, xreg=test_set$closing_price, h=length(test_indices))
  rmse_auto_arimax[i] <- sqrt(mean((test_set$top10_gross - auto_arimax_forecast$mean)^2))
}

# Calculate average RMSE across all folds
average_rmse_arimax <- mean(rmse_arimax)
average_rmse_auto_arimax <- mean(rmse_auto_arimax)

# Data frame for plotting
rmse_data <- data.frame(
  Model = rep(c("ARIMAX", "AUTO ARIMAX"), each = n_folds),
  RMSE = c(rmse_arimax, rmse_auto_arimax),
  Fold = rep(1:n_folds, 2)
)

# Plot the RMSE for both models
ggplot(rmse_data, aes(x = as.factor(Fold), y = RMSE, fill = Model)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_manual(values = c("#01BFC4", "#FB766D")) +
  labs(title = "RMSE for ARIMAX Models Across 5-Fold Cross-Validation", x = "Fold", y = "RMSE") +
  theme_minimal() +
  theme(legend.title = element_blank())

```
Both are giving the same RMSE/same model.

```{r}
summary(arimax_model)
```

Formula:

\[
(1 - \phi_1 B - \phi_2 B^2 - \phi_3 B^3 - \phi_4 B^4)(1 - B)Y_t = c + (1 + \theta_1 B)\varepsilon_t + \beta X_t
\]

where:
- \( B \) is the backshift operator.
- \( Y_t \) is the dependent variable (top10_gross) at time \( t \).
- \( \phi_1, \phi_2, \phi_3, \phi_4 \) are the coefficients for the autoregressive terms.
- \( \theta_1 \) is the coefficient for the moving average term.
- \( \varepsilon_t \) is the error term at time \( t \).
- \( X_t \) is the exogenous variable (closing_price) at time \( t \).
- \( c \) is the constant term (which is not provided in the summary but can be inferred or estimated separately).
- \( \beta \) is the coefficient for the exogenous variable.


```{r}
fimp<-forecast(arimax_model)

fxreg <- cbind(Exp = cp$mean,
              Imp = fimp$mean)

fcast <- forecast(fit, xreg=fxreg) #fimp$mean gives the forecasted values
autoplot(fcast) + xlab("Year") +
  ylab("GDP")
 
```


```{r}

 library(forecast)

# Define the number of periods you want to forecast
forecast_horizon <- 300

# Forecast 'closing_price' using the fitted ARIMA model for 'forecast_horizon' periods
fclosing <- forecast(CLOSING_fit, h=forecast_horizon)

# Forecast 'top10_gross' for 'forecast_horizon' periods using the forecasted 'closing_price'
fcast_top10 <- forecast(arimax_model, xreg=fclosing$mean, h=forecast_horizon)

# Plot the forecast of 'top10_gross'
plot(fcast_top10, main="Forecast of Top 10 Gross Revenue using ARIMAX Model")
lines(df$top10_gross, col="black")
legend("topleft", legend=c("Historical Top 10 Gross", "Forecasted Top 10 Gross"), col=c("black", "blue"), lty=1, bty="n")


```

